### 2023-06-01 doc/open-collab/selection/zenodo.Rmd

---
title: ""
output: html_document
---

```{r}
library(tidyverse)
library(here)
conflicted::conflict_prefer("filter", "dplyr")
zenodo_entries <- yaml::read_yaml(here("data-raw/zenodo.yaml"))
title_selection <- zenodo_entries %>% 
    map_dfr(~.x[c("title", "date", "doi", "keywords")]) 
View(title_selection)

remove_term_keep_collaborative <- function(variable, term) {
    !str_detect(variable, regex(term, ignore_case = TRUE)) &
        str_detect(variable, regex("collabora*", ignore_case = TRUE))
}

# TODO: Use ggcohort to track what is removed.
title_selection %>% 
    # Communication in the titles, e.g. "Scholarly communication" didn't fit 
    # with our aim.
    # filter(remove_term_keep_collaborative(title, "communication")) %>% 
    filter(str_detect(title, regex("communication", ignore_case = TRUE))) %>% 
    count()
    filter(str_detect(title, regex("publishing", ignore_case = TRUE))) %>% 
    View("Publishing")
    filter(str_detect(title, regex("open access", ignore_case = TRUE))) %>% 
    View("Open Access")
    View("")

title_selection %>% 
    filter(str_detect(title, regex("open science|collaboration", ignore_case = TRUE)))
    View()
```