### 2024-05-13 doc/review-stages.qmd
refactor: moved code from quarto into own function file

---
title: "Review stages"
format: html
---

```{r setup}
#| include: false
library(tidyverse)
library(here)
library(fs)
targets::tar_config_set(store = here("_targets"))
```

<!-- TODO: Include last date the articles were extracted from. -->

## Title review stage

During the first stage of the reviewing, we only looked at titles. Below
is the code to extract the titles we each individually reviewed, so we
can show some details about how many were included, excluded, and which
titles we agreed on or disagreed on.

```{r}
#| message: false
```

In the title screening stage, the number of titles we kept are listed in
@tbl-titles-kept.

```{r tbl-titles-kept}
#| message: false
#| tbl-cap: "Number of articles each reviewer kept after the first screening of the titles."
targets::tar_read(titles_kept) |>
  map(\(df) count(df, name = "Articles kept")) |>
  list_rbind(names_to = "Reviewers") |>
  # TODO: Include percent of total articles.
  knitr::kable()
```

We want to know how many of these titles we all agreed on between the
three reviewers:

```{r}
common_titles <- targets::tar_read(titles_agreed_on)
```

We seemed to agree on `r nrow(common_titles)` titles, with the full
title list shown in @tbl-titles-agreements.

```{r tbl-titles-agreements}
#| tbl-cap: !expr glue::glue("The {nrow(common_titles)} article titles we all agreed on and kept during the title screening stage.")
common_titles |>
  select("Article title" = title) |>
  knitr::kable()
```

One person will review the titles and flag any paper that might be
relevant but are not in agreement to keep.

```{r tbl-title-disagreements}
#| tbl-cap: "Article titles **not** kept between all reviewers during the title screening stage."
targets::tar_read(titles_disagreed_on) |>
  select("Article titles" = title) |>
  knitr::kable()
```

We saved a new csv file with all the disagreements and then made a copy
of that list in the file `resolved.csv`. This file contains some titles
that we disagreed on but that were briefly reviewed by one reviewer,
discussed, and included in the end.

The titles we agreed on were:

```{r}
#| tbl-cap: "Article titles we agreed to keep from the disagreement list."
targets::tar_read(titles_resolved_path) |>
  select("Article title" = title) |>
  knitr::kable()
```

## Abstract review stage

-   all will review the abstracts of the included papers
-   this will create our final list of included papers

## Reading stage

-   first skimming the papers and group by themes
-   then read a few from each theme to decide on what information to
    extract
-   then discuss together the themes to group by and information to
    extract
-   then we follow what we agreed on

## Writing stage

-   how to visualize?