### 2024-07-03 doc/review-stages.qmd
docs: :memo: update review stages file with what we kept, etc

---
title: "Review stages"
format: html
---

```{r setup}
#| include: false
library(tidyverse)
library(here)
library(fs)
targets::tar_config_set(store = here("_targets"))
source(here::here("R/abstract-review.R"))
```

<!-- TODO: Include last date the articles were extracted from. -->

## Title review stage

During the first stage of the reviewing, we only looked at titles. Below
is the code to extract the titles we each individually reviewed, so we
can show some details about how many were included, excluded, and which
titles we agreed on or disagreed on.

In the title screening stage, the number of titles we kept are listed in
@tbl-titles-kept.

```{r tbl-titles-kept}
#| message: false
#| tbl-cap: "Number of articles each reviewer kept after the first screening of the titles."
targets::tar_read(titles_kept) |>
  map(\(df) count(df, name = "Articles kept")) |>
  list_rbind(names_to = "Reviewers") |>
  # TODO: Include percent of total articles.
  knitr::kable()
```

We want to know how many of these titles we all agreed on between the
three reviewers:

```{r}
common_titles <- targets::tar_read(titles_agreed_on)
```

We seemed to agree on `r nrow(common_titles)` titles, with the full
title list shown in @tbl-titles-agreements.

```{r tbl-titles-agreements}
#| tbl-cap: !expr glue::glue("The {nrow(common_titles)} article titles we all agreed on and kept during the title screening stage.")
common_titles |>
  select("Article title" = title) |>
  knitr::kable()
```

One person will review the titles and flag any paper that might be
relevant but are not in agreement to keep.

```{r tbl-titles-disagreements}
#| tbl-cap: "Article titles **not** kept between all reviewers during the title screening stage."
targets::tar_read(titles_disagreed_on) |>
  select("Article titles" = title) |>
  knitr::kable()
```

We saved a new csv file with all the disagreements and then made a copy
of that list in the file `resolved.csv`. This file contains some titles
that we disagreed on but that were briefly reviewed by one reviewer,
discussed, and included in the end.

The titles we agreed on were:

```{r}
#| message: false
#| tbl-cap: "Article titles we agreed to keep from the disagreement list."
targets::tar_read(titles_resolved_path) |>
  read_csv() |>
  select("Article title" = title) |>
  knitr::kable()
```

The final list of titles we selected, between the agreed on ones as well
as the resolved ones, are:

```{r}
#| message: false
#| tbl-cap: "Article titles we reviewed and selected to continue for abstract review."
targets::tar_read(titles_selected) |>
  select("Article title" = title) |>
  knitr::kable()
```

## Abstract review stage

-   all will review the abstracts of the included papers
-   this will create our final list of included papers

The second phase of the review is looking over the abstracts. In
@tbl-abstracts-kept below are the number of articles each reviewer kept
after this second screening.

```{r tbl-abstracts-kept}
#| message: false
#| tbl-cap: "Number of articles each reviewer kept after the second screening of the articles."
targets::tar_read(abstracts_kept) |> 
  map(\(df) count(df, name = "Articles kept")) |>
  list_rbind(names_to = "Reviewers") |>
  # TODO: Include percent of total articles.
  knitr::kable()
```

Based on the abstract review, we want to know how many of these titles
we all agreed on between the three reviewers:

```{r}
common_abstracts <- targets::tar_read(abstracts_agreed_on)
```

We seemed to agree on `r nrow(common_abstracts)` abstracts, with the
full article list shown in @tbl-abstracts-agreements.

```{r tbl-abstracts-agreements}
#| tbl-cap: !expr glue::glue("The {nrow(common_abstracts)} article abstracts we all agreed on and kept during the abstract screening stage.")
targets::tar_read(abstracts_agreed_on) |>
  select("Article titles" = title) |>
  knitr::kable()
```

One person will review the titles and flag any paper that might be
relevant but are not in agreement to keep.

```{r tbl-abstracts-disagreements}
#| tbl-cap: "Article abstracts **not** kept between all reviewers during the title screening stage."
targets::tar_read(abstracts_disagreed_on) |>
  select("Article titles" = title) |>
  knitr::kable()
```

Like with the title review, we saved a new csv file with all the
disagreements and then made a copy of that list in the file
`resolved.csv`. This file contains some articles that we disagreed on
but that were briefly reviewed by one reviewer, discussed, and included
in the end.

The abstracts we agreed on were:

```{r}
#| message: false
#| tbl-cap: "Article abstracts we agreed to keep from the disagreement list."
targets::tar_read(abstracts_resolved_path) |>
  read_abstract_yaml() |>
  select("Article title" = title) |>
  knitr::kable()
```

The final list of titles we selected, between the agreed on ones as well
as the resolved ones, are:

```{r}
#| message: false
#| tbl-cap: "Article abstracts we reviewed and selected to continue to final article reviewing."
targets::tar_read(abstracts_selected) |>
  select("Article title" = title) |>
  knitr::kable()
```

## Reading stage

-   first skimming the papers and group by themes, writing a basic
    free-hand summary
-   then read a few papers from each theme to decide on what information
    to extract
-   then discuss together the themes to group by and information to
    extract

## Data extraction

-   we make a data extraction template
-   extract data and read the papers

## Writing stage

-   how to visualize?