### 2024-07-03 doc/reviews/daniel.qmd

---
title: 'Full Article Review'
---

# Open and reproducible neuroimaging: From study inception to publication

- File: data/review/fulltext/oa-id-W4295290221.pdf
- DOI: https://doi.org/10.1016/j.neuroimage.2022.119623
- OpenAlex ID: https://openalex.org/W4295290221

<!-- Write your notes down here -->

# Ten simple rules for helping newcomers become contributors to open projects

- File: data/review/fulltext/oa-id-W2972809368.pdf
- DOI: https://doi.org/10.1371/journal.pcbi.1007296
- OpenAlex ID: https://openalex.org/W2972809368

<!-- Write your notes down here -->

# Collaborative open science as a way to reproducibility and new insights in primate cognition research

- File: data/review/fulltext/oa-id-W4251805646.pdf
- DOI: https://doi.org/10.31234/osf.io/8w7zd
- OpenAlex ID: https://openalex.org/W4251805646

<!-- Write your notes down here -->

# Promoting FAIR Data Through Community-driven Agile Design: the Open Data Commons for Spinal Cord Injury (odc-sci.org)

- File: data/review/fulltext/oa-id-W3188722327.pdf
- DOI: https://doi.org/10.1007/s12021-021-09533-8
- OpenAlex ID: https://openalex.org/W3188722327

<!-- Write your notes down here -->

# UKRN Open Research Training Resources and Priorities Working Paper

- File: data/review/fulltext/oa-id-W4372403418.pdf
- DOI: https://doi.org/10.31219/osf.io/s2f6k
- OpenAlex ID: https://openalex.org/W4372403418

<!-- Write your notes down here -->

# Eleven Strategies for Making Reproducible Research and Open Science Training the Norm at Research Institutions

- File: data/review/fulltext/oa-id-W4378611187.pdf
- DOI: https://doi.org/10.31219/osf.io/kcvra
- OpenAlex ID: https://openalex.org/W4378611187

<!-- Write your notes down here -->

# Collaboration and Open Science Initiatives in Primate Research

- File: data/review/fulltext/oa-id-W3183365357.pdf
- DOI: https://doi.org/10.31219/osf.io/7c93a
- OpenAlex ID: https://openalex.org/W3183365357

<!-- Write your notes down here -->

# Open and reproducible practices in developmental psychology research: The workflow of the WomCogDev lab as an example

- File: data/review/fulltext/oa-id-W4229452124.pdf
- DOI: https://doi.org/10.31234/osf.io/73bwu
- OpenAlex ID: https://openalex.org/W4229452124

<!-- Write your notes down here -->

# Accelerating addiction research via Open Science and Team Science

- File: data/review/fulltext/oa-id-W4383376256.pdf
- DOI: https://doi.org/10.31234/osf.io/pbkrx
- OpenAlex ID: https://openalex.org/W4383376256

<!-- Write your notes down here -->

# Ten strategies to foster open science in psychology and beyond

- File: data/review/fulltext/oa-id-W4281886503.pdf
- DOI: https://doi.org/10.31234/osf.io/c38a2
- OpenAlex ID: https://openalex.org/W4281886503

<!-- Write your notes down here -->

# Open and collaborative tools for disaster management and risk reduction

- File: data/review/fulltext/oa-id-W4303191041.pdf
- DOI: https://doi.org/10.36335/vnjhm.2022(12).33-38
- OpenAlex ID: https://openalex.org/W4303191041

<!-- Write your notes down here -->

# UKRN ORCC Primer on Working in Open Research

- File: data/review/fulltext/oa-id-W4386723394.pdf
- DOI: https://doi.org/10.31219/osf.io/346hr
- OpenAlex ID: https://openalex.org/W4386723394

<!-- Write your notes down here -->

# Lessons Learned: A Neuroimaging Research Center’s Transition to Open and Reproducible Science

- File: data/review/fulltext/oa-id-W4283836446.pdf
- DOI: https://doi.org/10.31219/osf.io/fe74t
- OpenAlex ID: https://openalex.org/W4283836446

<!-- Write your notes down here -->