### 2024-08-15 doc/reviews/daniel.qmd
clean: the notes taken so far

---
title: 'Full Article Review'
---

# Open and reproducible neuroimaging: From study inception to publication

-   File: data/review/fulltext/oa-id-W4295290221.pdf
-   DOI: https://doi.org/10.1016/j.neuroimage.2022.119623
-   OpenAlex ID: https://openalex.org/W4295290221

## General themes

Tools to use.

## Type of paper

-   example
-   guide

## Other notes

This paper is not much about open collaboration but more about which
tools to use for open and reproducible research from start to finish of
a project.

# Ten simple rules for helping newcomers become contributors to open projects

-   File: data/review/fulltext/oa-id-W2972809368.pdf
-   DOI: https://doi.org/10.1371/journal.pcbi.1007296
-   OpenAlex ID: https://openalex.org/W2972809368

## General themes

-   community building
-   social rules
-   organization around the collaboration
-   new comers

## Type of paper

-   guide

## Other notes

-   Focus on building community: 1: be welcoming, 3. make governance
    explicit, 6. Develop forms of legitimate peripheral
    participation, 7. Make it easy for newcomers to get started.

# Collaborative open science as a way to reproducibility and new insights in primate cognition research

-   File: data/review/fulltext/oa-id-W4251805646.pdf
-   DOI: https://doi.org/10.31234/osf.io/8w7zd
-   OpenAlex ID: https://openalex.org/W4251805646

## General themes

-   practical setup of collaboration

## Type of paper

-   example

## Other notes

Focus on large-scale collaboration in psychology, because of a main
challenge of small samples sizes.

Project started in connection with a conference. First paper was a
pilot. Main aim of the project is to provide infrastructure for
latge-scale collaboration. Used mailing list to share information, then
slack for discussions. Use Google docs for writing and documentation and
github for code and data. Data analysis plans pre-registered at OSF.
Have a website - good for newcomers and sharing info. Topic for research
selected through voting. After organizing the data, they had a modelling
challenge. Also discuss challenges. Main one was how to make decisions.
Leadership team preferable. Also have author guidelines.

# Promoting FAIR Data Through Community-driven Agile Design: the Open Data Commons for Spinal Cord Injury (odc-sci.org)

-   File: data/review/fulltext/oa-id-W3188722327.pdf
-   DOI: https://doi.org/10.1007/s12021-021-09533-8
-   OpenAlex ID: https://openalex.org/W3188722327

## General themes

-   data sharing

## Type of paper

-   example

## Other notes

-   this paper is mostly about a database for sharing spinal injusry
    data. There are some points for how to improve sharing but not much
    about open collaboration
-   open data commons for spinal injury
-   Studies based on game theory suggest that data sharing might be
    beneficial if a collaborative approach is taken and data sharing is
    embraced as a community rather than by individuals
-   key events for the collaboration to be possible: - introduction of
    FAIR principles, - update from funders, funding given to specific
    projects
-   Open Data Commons for SCI (ODC-SCI, odc-sci.org), a platform to
    share tabular data of research in the field of spinal cord injury
-   Ain of the paper: to illustrate how members of research communities
    can work together toward the development of dedicated data sharing
    initiatives under the umbrella of FAIR
-   clearly highlighted the value of the massive demonstration/work for
    beta testing the site to reveal unforeseen problems
-   Moreover, participants pointed out the need for improving
    self-explanatory tutorials and help materials that would facilitate
    the learning experience for those who could not attend the workshop
-   "Registered users can request to become ODC-SCI Community members
    with further approval by the Leadership team." The most permissive
    account type is becoming an ODC-SCI Community member associated to
    an ODC-SCI lab, known as a Lab member

# UKRN Open Research Training Resources and Priorities Working Paper

-   File: data/review/fulltext/oa-id-W4372403418.pdf
-   DOI: https://doi.org/10.31219/osf.io/s2f6k
-   OpenAlex ID: https://openalex.org/W4372403418

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::

# Eleven Strategies for Making Reproducible Research and Open Science Training the Norm at Research Institutions

-   File: data/review/fulltext/oa-id-W4378611187.pdf
-   DOI: https://doi.org/10.31219/osf.io/kcvra
-   OpenAlex ID: https://openalex.org/W4378611187

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::

# Collaboration and Open Science Initiatives in Primate Research

-   File: data/review/fulltext/oa-id-W3183365357.pdf
-   DOI: https://doi.org/10.31219/osf.io/7c93a
-   OpenAlex ID: https://openalex.org/W3183365357

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::

# Open and reproducible practices in developmental psychology research: The workflow of the WomCogDev lab as an example

-   File: data/review/fulltext/oa-id-W4229452124.pdf
-   DOI: https://doi.org/10.31234/osf.io/73bwu
-   OpenAlex ID: https://openalex.org/W4229452124

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::

# Accelerating addiction research via Open Science and Team Science

-   File: data/review/fulltext/oa-id-W4383376256.pdf
-   DOI: https://doi.org/10.31234/osf.io/pbkrx
-   OpenAlex ID: https://openalex.org/W4383376256

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::

# Ten strategies to foster open science in psychology and beyond

-   File: data/review/fulltext/oa-id-W4281886503.pdf
-   DOI: https://doi.org/10.31234/osf.io/c38a2
-   OpenAlex ID: https://openalex.org/W4281886503

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::

# Open and collaborative tools for disaster management and risk reduction

-   File: data/review/fulltext/oa-id-W4303191041.pdf
-   DOI: https://doi.org/10.36335/vnjhm.2022(12).33-38
-   OpenAlex ID: https://openalex.org/W4303191041

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::

# UKRN ORCC Primer on Working in Open Research

-   File: data/review/fulltext/oa-id-W4386723394.pdf
-   DOI: https://doi.org/10.31219/osf.io/346hr
-   OpenAlex ID: https://openalex.org/W4386723394

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::

# Lessons Learned: A Neuroimaging Research Center's Transition to Open and Reproducible Science

-   File: data/review/fulltext/oa-id-W4283836446.pdf
-   DOI: https://doi.org/10.31219/osf.io/fe74t
-   OpenAlex ID: https://openalex.org/W4283836446

## General themes

::: content-hidden
Look for what the paper is generally about and some basic overall themes
and topics in the paper.
:::

## Type of paper

::: content-hidden
Is it a guide, community-building, an example of a workflow, etc.
:::

## Other notes

::: content-hidden
What we found was interesting or what we'd like to share, discuss, or
present in the main scoping review paper.
:::