### 2024-10-21 doc/reviews/themes.qmd
Made a file with identified themes with notes from individial papers

---
title: 'Analysis of themes'
---

# Tools

-   Open and reproducible neuroimaging: From study inception to
    publication

    -   study inception, planning, and ethics: OSF pre-registration,
        clinicaltrials.gov, aspredicted

    -   data processing and analysis: git, github, gitlab, etc

    -   many others

-   Collaborative open science as a way to reproducibility and new
    insights in primate cognition research

    -   Used mailing list to share information, then slack for
        discussions

    -   Use Google docs for writing and documentation

    -   Github for code and data

    -   Data analysis plans pre-registered at OSF

    -   Have a website - good for newcomers and sharing info

-   Open and reproducible practices in developmental psychology
    research: The workflow of the WomCogDev lab as an example

    -   Resources: lab meetings, slack, trello, OSF repository

-   Accelerating addiction research via Open Science and Team Science

    -   recommend sharing on OSF

    -   synthpop package in R creates a synthetic dataset with same
        statistical properties

-   Ten strategies to foster open science in psychology and beyond

    -   can get started by writing in markdown/quarto, using github

-   Lessons Learned: A Neuroimaging Research Center's Transition to Open
    and Reproducible Science

    -   use OSF for registering protocols

    -   focus on open software like R and Python, git and github

# Organization

-   Open and reproducible neuroimaging: From study inception to
    publication

    -   publishing Code of Conduct for collaborative projects is one
        practice that helps ensure a more welcoming and inclusive space
        for everyone regardless of background or identity

-   Ten simple rules for helping newcomers become contributors to open
    projects

    -   be welcoming

    -   make governance explicit

    -   develop forms of legitimate peripheral participation

    -   make it easy for newcomers to get started

-   Collaborative open science as a way to reproducibility and new
    insights in primate cognition research

    -   topic for research selected through voting

    -   also discuss challenges. Main one was how to make decisions

    -   leadership team preferable

    -   also have author guidelines.

-   Promoting FAIR Data Through Community-driven Agile Design: the Open
    Data Commons for Spinal Cord Injury (odc-sci.org)

    -   key events for the collaboration to be possible: - introduction
        of FAIR principles, update from funders, funding given to
        specific projects

    -   moreover, participants pointed out the need for improving
        self-explanatory tutorials and help materials that would
        facilitate the learning experience for those who could not
        attend the workshop

    -   registered users can request to become ODC-SCI Community members
        with further approval by the Leadership team.

-   Eleven Strategies for Making Reproducible Research and Open Science
    Training the Norm at Research Institutions

    -   perform replication or meta-research studies as course projects:
        Carefully define the scope of the project, Ensure that you have
        adequate support

    -   build communities: Foster accessible discussions (Consider
        running "beginner" and "advanced" community meetings)

    -   build communities: Organize regular meetings

-   Collaboration and Open Science Initiatives in Primate Research

    -   Agreeing on project selection, study design, stimuli, analysis
        plan, and findings' interpretation is not easy in big groups.
        (...) research questions are selected democratically - proposals
        for projects are submitted and members vote to select which
        projects will be carried out.

    -   Large-scale projects also have to develop authorship guidelines,
        which specify the minimal conditions that one person has to
        fulfill to qualify for authorship

    -   task forces within the project prepared different materials they
        then uploaded to github and pre-register

    -   Low entry barriers and the various ways in which a scientist can
        be involved in a large-scale collaboration (design, data
        collection, analysis, manuscript writing) offer a multitude of
        possibilities for researchers in different stages of their
        careers

-   Accelerating addiction research via Open Science and Team Science

    -   As part of this process, we made several decisions as a group
        (via an anonymous poll), such as which smallest effect size of
        interest to preregister and to which journal to submit the paper

-   Ten strategies to foster open science in psychology and beyond

    -   Participating in these large-scale projects is not easy and
        calls on researchers to develop specific standards and
        guidelines that ensure effective communication among
        collaborators and enable projects to develop coherently and
        cohesively

## Workflow

-   Open and reproducible practices in developmental psychology
    research: The workflow of the WomCogDev lab as an example

    -   They setup a clear structure for how to conduct the research -
        how a project is supposed to work

-   Lessons Learned: A Neuroimaging Research Center's Transition to Open
    and Reproducible Science

    -   also nice figure of workflow including open science practices in
        the workflow

# Training

-   UKRN Open Research Training Resources and Priorities Working Paper

    -   research cycle: planning, conducting, analysing, disseminating

    -   planning: team science guides, research co-production

    -   most resources focus on the planning and analysing stage

    -   very few resources for open collaboration, none were open and
        they were valued poorly compared to other practices

-   Eleven Strategies for Making Reproducible Research and Open Science
    Training the Norm at Research Institutions

    -   perform replication or meta-research studies as course projects:
        As the class collaborates on one project, participants also
        build skills for collaborative team science and gain experience
        leading small teams

# Barriers/limitations for implementation

-   Accelerating addiction research via Open Science and Team Science

    -   Barriers: within our current scientific culture, there are few
        incentives to work in large teams, all large-scale
        collaborations need leadership

# Levels of collaboration

## multi-center

-   Collaboration and Open Science Initiatives in Primate Research

## center

-   Lessons Learned: A Neuroimaging Research Center's Transition to Open
    and Reproducible Science

## lab

-   Open and reproducible practices in developmental psychology
    research: The workflow of the WomCogDev lab as an example