### 2024-10-21 doc/charting/Alessandroni-2022.qmd

---
title: 'Alessandroni-2022'
---

# Ten strategies to foster open science in psychology and beyond

-   File: data/review/fulltext/oa-id-W4281886503.pdf
-   DOI: https://doi.org/10.31234/osf.io/c38a2
-   OpenAlex ID: https://openalex.org/W4281886503

## Tools

### Specific tools mentioned - their function - where in the researh process used

-   Markdown/quarto - collaborative writing - all
-   Github - version control - all

## Organizational structure for open collaboration

### Governance

-   use open tools for collaboration
-   develop specific standards and guidelines that ensure effective
    communication among collaborators and enable projects to develop
    coherently and cohesively

## Educational perspectives

### Educational needs

-   Introduce open science practices into courses
-   Institutions can create courses in open science practices

## Barriers

### Barriers for open science

-   Incorporating open tools may have a steep learning curve
-   Open science in the curricula vitae of academics must become a
    requirement and not just a nice addition
-   The problem with current incentive structures such as impact factor