### 2024-10-21 doc/charting/Bush-2022.qmd

---
title: 'Bush-2022'
---

# Lessons Learned: A Neuroimaging Research Center's Transition to Open and Reproducible Science

-   File: data/review/fulltext/oa-id-W4283836446.pdf
-   DOI: https://doi.org/10.31219/osf.io/fe74t
-   OpenAlex ID: https://openalex.org/W4283836446

## Tools

### Specific tools mentioned - their function - where in the researh process used

-   Github - open-source code, version control, collaboration -
    analysing
-   Python - open-source programming - analysing
-   R - open-source programming - analysing
-   arXiv/bioRxiv/psyRxiv - preprint sharing - disseminating
-   OpenNeuro - data sharing - disseminating
-   Open Science Framework - preregistration - planning
-   clinicaltrials.gov - preregistration - planning

## Organizational structure for open collaboration

### Governance

### Workflow

-   hypothesis - experiment - data collection - data analysis -
    reporting

## Educational perspectives

### Educational needs

-   people in the organization need to learn additional programming,
    version control, and data management skills and work with unfamiliar
    naming conventions and directory structures mandated by the used
    standards

## Barriers

### Barriers for open science

-   costs of effort to be borne by individuals already engaged in
    challenging, time-consuming work
-   during transition, the research productivity will suffer