### 2024-10-21 doc/charting/Dora-2024.qmd

---
title: 'Dora-2024'
---

# Accelerating addiction research via Open Science and Team Science

-   File: data/review/fulltext/oa-id-W4383376256.pdf
-   DOI: https://doi.org/10.31234/osf.io/pbkrx
-   OpenAlex ID: https://openalex.org/W4383376256

## Tools

### Specific tools mentioned - their function - where in the researh process used

::: content-hidden
For instance: "OSF - pre-registration - planning". You can mention
multiple tools. Just make another bullet.

The research process is: planning, conducting, analysing, disseminating.
If a combination write several, if all write "all"
:::

-

## Organizational structure for open collaboration

### Governance

::: content-hidden
For instance: "Develop forms of legitimate peripheral participation."

These are bit of information and recommendations for governance of the
project.
:::

-   Used anonymous polls to make critical decisions in a large
    collaborative project
-   Important to setup a structure for leadership of the project and
    that the leadership is held accountable to decisions made
-   setup a structure to regularly ask for formal and informal feedback
    from members

### Workflow

::: content-hidden
For instance: "designing the project - preregistration - creating the
experiment - recruitment and data collection - data analysis -
dissemination".

These are overall process highlights
:::

-

## Educational perspectives

### Educational needs

::: content-hidden
For instance: "few resources for open collaboration, none were open and
they were valued poorly compared to other practices"

What does the paper write is the need for education in the field?
:::

-   Education in open science needed to increase uptake

## Barriers

### Barriers for open science

::: content-hidden
For instance: "lack of education"

Which barriers do the paper indicate for implementation of open science?
:::

-   no open science standards in journals
-   few opportunities for registered reports
-   there are few incentives to collaborate in large teams
-   grants are often given to an individual not a team