### 2024-10-21 doc/charting/Kohrs-2023.qmd
Add files via upload

---
title: 'Khors-2023'
---

# Eleven Strategies for Making Reproducible Research and Open Science Training the Norm at Research Institutions

-   File: data/review/fulltext/
-   DOI: https://doi.org/10.31219/osf.io/kcvra
-   OpenAlex ID: W4378611187

## Tools

### Specific tools mentioned - their function - where in the researh process used

-   Virtual unconference - not a tool per se, but a nice way of networking and sharing ideas - can be used either early to brainstorm and network or later to disseminate ideas
-   COARA - fairer way of assessing research output than impact factors - Not used, just mentioned.
-   Open Science Labs - not a tool per se, but dedicated staff and organizations who works solely to contribute and assist on designing open science - study design, but they can     assist in other stages. other decentralized hubs exist such as FORRT or STEM education hub.
-   SCOSS - application for funding to be able to publish with open science standards - publication.

## Organizational structure for open collaboration

### Systematic changes to promote Open Science

-   Make open science necessary for the curriculum in universities (Munich university as example) and add courses that touches on openness, preregistration, etc.
-   Thesis are required to have open science elements to them. including collaborative thesis
-   Open Science requirements in the curriculum needed to hire individuals.

## Educational perspectives

-   Here education is the central aspect of the structure to build open science the norm in an institution
-   They suggest offering open science courses, but also integreate open science in other courses
-   They also suggest that hand-on training is the best way to learn, some ways of doing this is by making courses more interactive or have research groups go on open science        courses. Some of the courses or examples can be replications or collaborative thesis.
-   An essential part of education is allowing for spaces where researchers can work, such as journal clubs, resource hubs or hackathons.

## Barriers

### Barriers for open science

-   Allocation for resources.
-   In general all the things required to change expressed in the topics above: open science not being part of the standard curriculum of universities or a required skill for        when hiring. The lack of spaces and infrastructure to network and collaborate. Innovative papers being preferred by mainstream media than replications.