### 2024-10-22 doc/charting/Dora-2024.qmd
update data extraction from my studies

---
title: 'Dora-2024'
---

# Accelerating addiction research via Open Science and Team Science

-   File: data/review/fulltext/oa-id-W4383376256.pdf
-   DOI: https://doi.org/10.31234/osf.io/pbkrx
-   OpenAlex ID: https://openalex.org/W4383376256

## Characteristics of the paper

-   Type of paper (e.g., tips, example): guide
-   Themes (e.g., tools, organization): organization
-   Other keywords (e.g., newcomers): team science, individual-centered

## Organizational structure for open collaboration

### Governance

-   Used anonymous polls to make critical decisions in a large
    collaborative project
-   Important to setup a structure for leadership of the project and
    that the leadership is held accountable to decisions made
-   setup a structure to regularly ask for formal and informal feedback
    from members

## Educational perspectives

### Educational needs

-   Education in open science needed to increase uptake

## Barriers

### Barriers for open science

-   no open science standards in journals
-   few opportunities for registered reports
-   there are few incentives to collaborate in large teams
-   grants are often given to an individual not a team