### 2024-10-22 doc/charting/Turoman-2022.qmd

---
title: 'Turoman-2022'
---

# Open and reproducible practices in developmental psychology research: The workflow of the WomCogDev lab as an example

-   File: data/review/fulltext/oa-id-W4229452124.pdf
-   DOI: https://doi.org/10.31234/osf.io/73bwu
-   OpenAlex ID: https://openalex.org/W4229452124

## Characteristics of the paper

-   Type of paper (e.g., tips, example): example
-   Themes (e.g., tools, organization): tools, workflow
-   Other keywords (e.g., newcomers): lab-centered

## Tools

### Specific tools mentioned - their function - where in the researh process used

-   Slack - communication - all
-   Trello - project management - planning
-   Open Science Framework - repository - planning, analysing,
    dissemination
-   R - data analysis - analysing
-   Jasp - data analysis - analysing
-   Zenodo - data sharing - dissemination
-   OpenNeuro - data sharing - dissemination

## Organizational structure for open collaboration

### Governance

-   Setup a clear structure for how to conduct the research considering
    open science in each step
-   Build in steps for error correction

### Workflow

-   designing the project - preregistration - creating the experiment -
    recruiting and data collection - data analysis - dissemination

## Barriers

### Barriers for open science

-   Not everyone have the same prerequisites. There are conscious
    tradeoffs between trying to be as open as possible, and the
    feasibility of applying an open practice
-   Some practices can take additional time
-   For the workflow to work, lab members must follow it quite closely
-   The learning curve can be steep, especially for people who arrive
    from more goal-oriented labs