### 2024-10-22 doc/charting/ManyPrimates-2021.qmd
added a few points to this paper

---
title: 'ManyPrimates-2021'
---

# Collaborative open science as a way to reproducibility and new insights in primate cognition research

-   File: data/review/fulltext/
-   DOI: https://doi.org/10.31234/osf.io/8w7zd
-   OpenAlex ID: W4251805646

## Tools

### Specific tools mentioned - their function - where in the researh process used

-   Github - version control - all
-   Registered reports - help designing analysis to prove specific
    hypothesis - planning
-   p-curve software - helps avoiding overrepresantation of positive
    results - analysis
-   A.P.E.S Wiki - open access platform for standardized research and
    conservation of the data - dissemination
-   PRIMatE Research Exchange - platform to exchange data, ideas and
    protocols - dissemination

## Organizational structure for open collaboration

### Setting up a network of collaborators

-   ManyPrimates wanted a lasting infrastructure
-   They developed a mailing list where researchers can exchange their
    ideas and analysis.
-   They developed a twitter (now X) account to disseminate and extend
    their work to other researchers.

### Setting up the bases for collaboration and a test example to educate people in open sciences

-   They developed ethical guidelines for all collaborators
    (non-invasive research, always open science, author list)
-   Research questions are selected democratically - proposals for
    projects are submitted and members vote to select which projects
    will be carried out
-   They ran a pilot study that tried replicating analysis on short-term
    memory with more sample size.
-   They preregistered protocols, created code and hosted data in
    repositories so that other coworkers can run the analysis
    systematically.
-   Data was collected.
-   Other workforce merged the data and visualized the results.
-   With this pipeline, they decided to run other analysis. This
    highlights the importance of a good test study.
-   Task forces within the project prepared different materials they
    then uploaded to github and pre-register

## Educational perspectives

-   Large-scale collaboration are a great opportunity for making more
    researchers get in touch with open science practices.
-   Creating websites like A.P.E.S Wiki or a mailing list creates a
    network of researchers that can follow similar practices.
-   Replication studies like the mentioned MP1 project on short-term
    memory can be of great educational value and help set up and
    infrastructure to promote openness.

## Barriers

### Barriers for open science

-   Pre-registration might neglect innovation due to their rigid
    structure, but creates very solid results.