### 2024-10-22 doc/charting/Kohrs-2023.qmd
updated list of tools to sync with format in other papers and added info to other sections

---
title: 'Khors-2023'
---

# Eleven Strategies for Making Reproducible Research and Open Science Training the Norm at Research Institutions

-   File: data/review/fulltext/
-   DOI: https://doi.org/10.31219/osf.io/kcvra
-   OpenAlex ID: W4378611187

## Tools

### Specific tools mentioned - their function - where in the researh process used

-   Virtual conference - networking and sharing ideas - all
-   COARA - fairer way of assessing research output than impact
    factors - dissemination
-   SCOSS - application for funding to be able to publish with open
    science standards - dissemination

## Organizational structure for open collaboration

### Systematic changes to promote Open Science

-   Make open science necessary for the curriculum in universities
    (Munich university as example) and add courses that touches on
    openness, preregistration, etc.
-   Thesis are required to have open science elements to them. Including
    collaborative thesis
-   Open Science requirements in the CV needed to hire individuals
-   Carefully define the scope of the project, Ensure that you have
    adequate support
-   Foster accessible discussions (Consider running "beginner" and
    "advanced" community meetings)
-   Build communities:Organize regular meetings

## Educational perspectives

-   Here education is the central aspect of the structure to build open
    science the norm in an institution
-   Offering open science courses, but also integreate open science in
    other courses
-   Hands-on training is the best way to learn, some ways of doing this
    is by making courses more interactive or have research groups go on
    open science courses. Some of the courses or examples can be
    replications or collaborative thesis.
-   Perform replication or meta-research studies as course projects
-   An essential part of education is allowing for spaces where
    researchers can work, such as journal clubs, resource hubs or
    hackathons.

## Barriers

### Barriers for open science

-   Allocation for resources.
-   In general all the things required to change expressed in the topics
    above: open science not being part of the standard curriculum of
    universities or a required skill for when hiring. The lack of spaces
    and infrastructure to network and collaborate. Innovative papers
    being preferred by mainstream media than replications.