### 2024-10-22 doc/scispace/README.Rmd
scispace charting of the papers with documentation

# Description of process and instructions for SciSpace

We used SciSpace to additionally perform data charting using the themes
we identified.

All papers were uploaded to SciSpace.

## Instructions

The summary columns were as follows with the following instructions:

### TL_DR

TL;DR column used the default setting for TL;DR summary from SciSpace.

### Tools

A custom column.

Named: tools suggested for open collaboration

Specification: what are the specific tools mentioned in the paper to
help researchers to open and collaborative science? For instance open
science framework (OSF) or github.

### Organization

A custom column.

Named: organization of open collaboration

Specification: what are the suggested practices for open collaboration?
this also includes own examples in the papers. This could be the use of
slack for communication and github for coding.

### Barriers

A custom column.

Named: barrier

Specification: what are the barriers for doing more open science
mentioned in the paper?