### 2024-10-23 doc/charting/Alessandroni-2022.qmd
Update Alessandroni-2022.qmd
Added a couple of barriers I found relevant. Other than that, this paper is all over the place and it has less workflow or structure than otherrs. Interesting, but does not add anything new if you have read the others.

---
title: 'Alessandroni-2022'
---

# Ten strategies to foster open science in psychology and beyond

-   File: data/review/fulltext/oa-id-W4281886503.pdf
-   DOI: https://doi.org/10.31234/osf.io/c38a2
-   OpenAlex ID: https://openalex.org/W4281886503

## Characteristics of the paper

-   Type of paper (e.g., tips, example): guide
-   Themes (e.g., tools, organization): organization, tools
-   Other keywords (e.g., newcomers):

## Tools

### Specific tools mentioned - their function - where in the researh process used

-   Markdown/quarto - collaborative writing - all
-   Github/Gitlab - version control - all

## Organizational structure for open collaboration

### Governance

-   use open tools for collaboration
-   develop specific standards and guidelines that ensure effective
    communication among collaborators and enable projects to develop
    coherently and cohesively

## Educational perspectives

### Educational needs

-   Introduce open science practices into courses
-   Institutions can create courses in open science practices

## Barriers

### Barriers for open science

-   Incorporating open tools may have a steep learning curve
-   Concerns about data safety
-   High prices for making papers open
-   Open science in the curricula vitae of academics must become a
    requirement and not just a nice addition
-   The problem with current incentive structures such as impact factor