### 2024-10-26 doc/charting/_template.qmd
docs: :pencil2: very minor edits to template, most format and one typo

---
title: 'Charting form template'
---

::: content-hidden
Fill in one form per paper. Name the form according to first author and
year like this: "Turoman-2022". Copy-paste the information from the
full-text review below.
:::

# Title of publication

-   File:
-   DOI:
-   OpenAlex ID:

## Characteristics of the paper

-   Type of paper (e.g., tips, example):
-   Themes (e.g., tools, organization):
-   Other keywords (e.g., newcomers):

## Tools

### Specific tools mentioned; their function; where in the research process used

::: content-hidden
For instance: "OSF; pre-registration; planning". You can mention
multiple tools. Just make another bullet.

The research process is: planning, conducting, analysing, disseminating.
If a combination write several, if all write "all"
:::

-

## Organizational structure for open collaboration

### Governance

::: content-hidden
For instance: "Develop forms of legitimate peripheral participation."

These are bit of information and recommendations for governance of the
project.
:::

-

### Workflow

::: content-hidden
For instance: "designing the project - preregistration - creating the
experiment - recruitment and data collection - data analysis -
dissemination".

These are overall process highlights
:::

-

## Educational perspectives

### Educational needs

::: content-hidden
For instance: "few resources for open collaboration, none were open and
they were valued poorly compared to other practices"

What does the paper write is the need for education in the field?
:::

-

## Barriers

### Barriers for open science

::: content-hidden
For instance: "lack of education"

Which barriers do the paper indicate for implementation of open science?
:::

-