### 2024-10-26 doc/charting/TorresEspin-2021.qmd

---
title: 'TorresEspin-2021'
---

# Promoting FAIR Data Through Community-driven Agile Design: the Open Data Commons for Spinal Cord Injury (odc-sci.org)

-   File: data/review/fulltext/oa-id-W3188722327.pdf
-   DOI: https://doi.org/10.1007/s12021-021-09533-8
-   OpenAlex ID: https://openalex.org/W3188722327

## Characteristics of the paper

-   Type of paper (e.g., tips, example):
    -   Case study
    -   Example workflow
-   Themes (e.g., tools, organization):
    -   tools
    -   organization
-   Other keywords (e.g., newcomers):
    -   Data infrastructure
    -   FAIR data
    -   Case study
    -   Agile design
    -   Templates
    -   Examples
    -   Strategies
    -   Software development

## Tools

### Specific tools mentioned; their function; where in the research process used

::: content-hidden
For instance: "OSF; pre-registration; planning". You can mention
multiple tools. Just make another bullet.

The research process is: planning, conducting, analysing, disseminating.
If a combination write several, if all write "all"
:::

-   GitHub?; Seems to use GitHub to at least build the website;
    dissemination?
-   Agile development; iteratively develop project and help with project
    management; workflow, planning, development
-   Staged development; distinct stages mixed with iterations; planning,
    development

## Organizational structure for open collaboration

### Governance

::: content-hidden
For instance: "Develop forms of legitimate peripheral participation."

These are bit of information and recommendations for governance of the
project.
:::

-   Building and encouraging community approval, support, and ownership
    via workshop events done in collaboration with major
    organizations/conferences.
    -   Engaging multiple levels of stakeholders
    -   Collaborate with funding agencies early on
-   Split into different teams, which have positions that aim to be
    about 3 year to rotate around people:
    -   Leadership board to coordinate the development and operation
    -   Executive board for oversight and be involved in executive
        decisions
    -   Community board to engage in community, get feedback from
        workshops
    -   Data science team for data curation, quality control, and
        revision

### Workflow

::: content-hidden
For instance: "designing the project - preregistration - creating the
experiment - recruitment and data collection - data analysis -
dissemination".

These are overall process highlights
:::

-   Took a multi-staged approach to establishing the group/community.

    -   At each stage, it was slowly about introducing the concepts to a
        broader and broader audience and getting feedback all along the
        way.

-   Development follows principles of agile software development by
    getting requirements from users, designing and developing those
    requirements, seeking feedback from users on what was developed, and
    testing the developed features.

-   Use of the iterative development helped them identify and respond to
    issues that came up. This is something that could be something that
    non-software projects could use to help with collaboration and
    maintaining momentum and motivation.

## Educational perspectives

### Educational needs

::: content-hidden
For instance: "few resources for open collaboration, none were open and
they were valued poorly compared to other practices"

What does the paper write is the need for education in the field?
:::

-   No education or training was mentioned in this paper.

## Barriers

### Barriers for open science

::: content-hidden
For instance: "lack of education"

Which barriers do the paper indicate for implementation of open science?
:::

-   Building up an open collaboration project takes years of continued
    work and effort.

-   While the project's output is a data sharing platform, the process
    to get there required a high level of collaboration.

-   Required a lot of expertise not typically found in researchers (e.g.
    software development, UI/UX design, data engineering, writing
    user-friendly documentation/tutorials).

-   Required getting regular direct feedback from potential users of
    project, which wasn't always easy.