### 2024-10-26 doc/charting/Sholler-2019.qmd
docs: :memo: finished charting sholler 2019

---
title: Sholler-2019
---

This paper seems more targeted to general open projects and not strictly
about open collaboration within research. However, it is useful
information contained within on some general steps to take to encourage
and continue collaboration in open projects and in as open ways as
possible.

# Ten simple rules for helping newcomers become contributors to open projects

-   File: data/review/fulltext/oa-id-W2972809368.pdf
-   DOI: https://doi.org/10.1371/journal.pcbi.1007296
-   OpenAlex ID: https://openalex.org/W2972809368

## Characteristics of the paper

-   Type of paper (e.g., tips, example): guide
-   Themes (e.g., tools, organization):
    -   community-building
-   Other keywords (e.g., newcomers):
    -   Onboarding
    -   Documentation
    -   Contributing
    -   Centralization
    -   Knowledge management
    -   Communication
    -   Guidelines

## Tools

### Specific tools mentioned; their function; where in the research process used

-   Not relevant.

## Organizational structure for open collaboration

### Governance

-   Make governance explicit (e.g. have clear documents describing who
    are in leadership positions, who makes decisions, and how decisions
    are made).
-   Include forms of legitimate peripheral participation that newcomers
    are most likely to start doing, such as commenting on issues or
    discussions or creating issues when they find something that needs
    attention.

### Workflow

-   There wasn't anything specific about workflows in this paper.

## Educational perspectives

### Educational needs

-   Ensure there are clear contributing guidelines and code of conduct
    to help contributors out.
-   Make it easy for people to get started, for instance, clear
    instructions for newcomers on how to collaborate.

## Barriers

### Barriers for open science

-   There wasn't anything specific about barriers in this paper.