### 2025-01-08 doc/charting/ManyPrimates-2019.qmd

---
title: ManyPrimates-2019
---

::: callout-note
This paper wasn't the most relevant, as it didn't go into much detail on
collaboration.
:::

# Collaborative open science as a way to reproducibility and new insights in primate cognition research

-   File: data/review/fulltext/oa-id-W4251805646.pdf
-   DOI: https://doi.org/10.31234/osf.io/8w7zd
-   OpenAlex ID: https://openalex.org/W4251805646

## Characteristics of the paper

-   Type of paper (e.g., tips, example): review?, example
-   Themes (e.g., tools, organization):
    -   organization
    -   tools
    -   review
-   Other keywords (e.g., newcomers):
    -   Systematic review
    -   Large-scale collaboration
    -   Large-scale infrastructure

## Tools

### Specific tools mentioned; their function; where in the research process used

::: content-hidden
For instance: "OSF; pre-registration; planning". You can mention
multiple tools. Just make another bullet.

The research process is: planning, conducting, analysing, disseminating.
If a combination write several, if all write "all"
:::

-   GitHub; sharing materials and final papers; archiving, dissemination
-   Mailing lists; communication; all
-   Google Docs; collaborative writing; writing, planning
-   Google Drive; sharing of procedures and documentation; all
-   Slack
-   OSF; archiving protocols and analysis plans; archiving
-   Organization website; share news and communicate updates; all

## Organizational structure for open collaboration

### Governance

::: content-hidden
For instance: "Develop forms of legitimate peripheral participation."

These are bit of information and recommendations for governance of the
project.
:::

-   Started from symposium as part of regular major conference.
-   Established a mailing list to connect everyone.
-   No clear description of how they organized their governance.

### Workflow

::: content-hidden
For instance: "designing the project - preregistration - creating the
experiment - recruitment and data collection - data analysis -
dissemination".

These are overall process highlights
:::

-   They weren't completely clear on how exactly they used the workflows
    they (very briefly) describe.
-   Use GitHub or Google Drive to share things.
-   Use Google Docs to write papers.
-   Use their website to disseminate news and findings.
-   Projects should preregister the plans, and all data, code, and
    materials are uploaded to GitHub.

## Educational perspectives

### Educational needs

::: content-hidden
For instance: "few resources for open collaboration, none were open and
they were valued poorly compared to other practices"

What does the paper write is the need for education in the field?
:::

-   No major educational comments were made.
-   Outreach of results and activities is done through social media and
    conferences.

## Barriers

### Barriers for open science

::: content-hidden
For instance: "lack of education"

Which barriers do the paper indicate for implementation of open science?
:::

-   Logistical. Impossible to use the exact same methods across all
    studies, so re-using of methods and materials wasn't very directly
    useful.
-   Also technical and knowledge, not enough funding to hire relevant
    expertise and technical resources.
-   Lacked a clear governance structure, which is something they want to
    do in the future.