### 2025-01-08 doc/charting/index.qmd
docs: :art: mostly formatting of docs to fit website

---
title: Charting plan
---

This is a short description of the charting plan.

1.  Each author is assigned specific papers to chart using the charting
    template
2.  The other authors reviews the charted form and updates if necessary
3.  If there are any disagreements, these will be resolved by discussion
4.  In the end, the data will be transferred to a data table

## Papers to chart

### Daniel

-   Lessons Learned: A Neuroimaging Research Center's Transition to Open
    and Reproducible Science
-   Ten strategies to foster open science in psychology and beyond
-   Accelerating addiction research via Open Science and Team Science
-   Open and reproducible practices in developmental psychology
    research: The workflow of the WomCogDev lab as an example

### Mario

-   Collaboration and Open Science Initiatives in Primate Research
-   Eleven Strategies for Making Reproducible Research and Open Science
    Training the Norm at Research Institutions
-   UKRN Open Research Training Resources and Priorities Working Paper
-   Open and reproducible neuroimaging: From study inception to
    publication

### Luke

-   Promoting FAIR Data Through Community-driven Agile Design: the Open
    Data Commons for Spinal Cord Injury (odc-sci.org)
-   Collaborative open science as a way to reproducibility and new
    insights in primate cognition research
-   Ten simple rules for helping newcomers become contributors to open
    projects

## Template to follow

::: {.callout-note appearance="minimal"}
{{< include _template.qmd >}}
:::