### 2025-01-08 doc/appendices.qmd

---
title: Appendices
---

This section lists additional files that we used or created in the
process of doing this scoping review project.

-   [Instructions](instructions/index.qmd): This docume has instructions
    for doing a scoping review as well as the PRISMA-ScR checklist.
-   [Full-text reviews](reviews/index.qmd): These documents contain each
    author's review of the full-text documents we selected after the
    title and abstract screening, as well as some grouping of them into
    themes and a listing of the titles of the documents we selected at
    each phase of the project.
-   [Charting](charting/index.qmd): This contains the structured
    charting documents we used for each full-text document we selected.
-   [Reflections](reflections/index.qmd): These are our informal
    reflections we wrote in to help us identify challenges and ways we
    improved on that.