### 2025-01-08 doc/charting/index.qmd

---
title: Charting plan
---

This is a short description of the charting plan.

1.  Each author is assigned specific papers to chart using the charting
    template
2.  The other authors reviews the charted form and updates if necessary
3.  If there are any disagreements, these will be resolved by discussion
4.  In the end, the data will be transferred to a data table

## Papers to chart

### Daniel

-   [Bush-2022](Bush-2022.qmd): Lessons Learned: A Neuroimaging Research
    Center's Transition to Open and Reproducible Science
-   [Alessandroni-2022](Allessandroni-2022.qmd): Ten strategies to
    foster open science in psychology and beyond
-   [Dora-2024](Dora-2024.qmd): Accelerating addiction research via Open
    Science and Team Science
-   [Turoman-2022](Turoman-2022.qmd): Open and reproducible practices in
    developmental psychology research: The workflow of the WomCogDev lab
    as an example

### Mario

-   [ManyPrimates-2021](ManyPrimates-2021.qmd): Collaboration and Open
    Science Initiatives in Primate Research
-   [Kohrs-2023](Kohrs-2023.qmd): Eleven Strategies for Making
    Reproducible Research and Open Science Training the Norm at Research
    Institutions
-   [Grange-2023](Grange-2023.qmd): UKRN Open Research Training
    Resources and Priorities Working Paper
-   [Niso-2022](Niso-2022.qmd): Open and reproducible neuroimaging: From
    study inception to publication

### Luke

-   [TorresEspin-2021](TorresEspin-2021.qmd): Promoting FAIR Data
    Through Community-driven Agile Design: the Open Data Commons for
    Spinal Cord Injury (odc-sci.org)
-   [ManyPrimates-2019](ManyPrimates-2019.qmd): Collaborative open
    science as a way to reproducibility and new insights in primate
    cognition research
-   [Scholler-2019](Scholler-2019.qmd): Ten simple rules for helping
    newcomers become contributors to open projects

## Template to follow

::: {.callout-note appearance="minimal"}
{{< include _template.qmd >}}
:::